Polarized Cubical Types

### 2014-05-22 ring.tex
thoughts

\newcommand{\bdim}[1]{\textsf{dim}(#1)}
\newcommand{\Rule}[4]{
\makebox{{\rm #1}
$\displaystyle
\frac{\begin{array}{l}#2\\\end{array}}
{\begin{array}{l}#3\\\end{array}}$
 #4}}
\newcommand{\proves}{\vdash}
\newcommand{\symc}[1]{\mathit{sym}~#1}
\newcommand{\jdg}[3]{#2 \proves_{#1} #3}
\newcommand{\adjoint}[1]{#1^{\dagger}}
\newcommand{\iso}{\leftrightarrow}
\newcommand{\identlp}{\mathit{identl}_+}
\newcommand{\identrp}{\mathit{identr}_+}
\newcommand{\swapp}{\mathit{swap}_+}
\newcommand{\assoclp}{\mathit{assocl}_+}
\newcommand{\assocrp}{\mathit{assocr}_+}
\newcommand{\identlt}{\mathit{identl}_*}
\newcommand{\identrt}{\mathit{identr}_*}
\newcommand{\swapt}{\mathit{swap}_*}
\newcommand{\assoclt}{\mathit{assocl}_*}
\newcommand{\assocrt}{\mathit{assocr}_*}
\newcommand{\distz}{\mathit{dist}_0}
\newcommand{\factorz}{\mathit{factor}_0}
\newcommand{\dist}{\mathit{dist}}
\newcommand{\factor}{\mathit{factor}}
\newcommand{\idc}{\mathit{id}}
\newcommand{\plus}{\raisebox{.4\height}{\scalebox{.6}{+}}}
\newcommand{\minus}{\raisebox{.4\height}{\scalebox{.8}{-}}}
\newcommand{\mm}{\texttt{-}}
\newcommand{\pp}{\texttt{+}}
\newcommand{\negp}[1]{\textit{neg}(#1)}
\newcommand{\inl}[1]{\textsf{inl}~#1}
\newcommand{\inr}[1]{\textsf{inr}~#1}
\newcommand{\lolli}{\multimap} 
\newcommand{\cubt}{\mathbb{T}}
\newcommand{\den}[1]{\llbracket #1 \rrbracket}
\newcommand{\nodet}[2]{\fcolorbox{black}{white}{$#1$}\fcolorbox{black}{gray!20}{$#2$}}

@@

\begin{figure*}
\[\begin{array}{c}
\nodet{S_1}{S_2}
\quad\otimes\quad
\nodet{(\nodet{S_3}{S_4})}{(\nodet{S_5}{S_6})} \quad= \\
\\
\nodet{(\nodet{{(\nodet{S_1 \times S_3}{S_1 \times S_4})}}
              {{(\nodet{S_1 \times S_5}{S_1 \times S_6})}})}
      {(\nodet{{(\nodet{S_2 \times S_3}{S_2 \times S_4})}}
              {{(\nodet{S_2 \times S_5}{S_2 \times S_6})}})}
\end{array}\]
\caption{\label{mult}Example of multiplication of two cubical sets.}
\end{figure*}

\begin{figure*}
\begin{tikzpicture}
\node[above] at (0,0) {\pp};
\node[left] at (0,0) {$S_1$};
\draw[fill] (0,0) circle [radius=0.05];
\node[above] at (0.6,0) {\mm};
\node[right] at (0.6,0) {$S_2$};
\draw[fill] (0.6,0) circle [radius=0.05];
\draw[-] (0,0) -- (0.6,0);
\node at (1.6,0) {$\otimes$};

%%
\node[left] at (2.5,-0.5) {\pp\pp};
\node[below] at (2.5,-0.5) {$S_3$};
\draw[fill] (2.5,-0.5) circle [radius=0.05];
\node[right] at (3.5,-0.5) {\pp\mm};
\node[below] at (3.5,-0.5) {$S_4$};
\draw[fill] (3.5,-0.5) circle [radius=0.05];
\draw[-] (2.5,-0.5) -- (3.5,-0.5);
\draw[-] (2.5,-0.5) -- (2.5,0.5);
\node[left] at (2.5,0.5) {\mm\pp};
\node[above] at (2.5,0.5) {$S_5$};
\draw[fill] (2.5,0.5) circle [radius=0.05];
\node[right] at (3.5,0.5) {\mm\mm};
\node[above] at (3.5,0.5) {$S_6$};
\draw[fill] (3.5,0.5) circle [radius=0.05];
\draw[-] (2.5,0.5) -- (3.5,0.5);
\draw[-] (3.5,-0.5) -- (3.5,0.5);
%% 
\node at (5,0) {$=$};
%% 
\node[left] at (7.5,0.75) {$(S_2 \times S_3)$\,\mm\pp\pp};
\draw[fill] (7.5,0.75) circle [radius=0.05];
\node[right] at (9.5,0.75) {$(S_2 \times S_4)$\,\mm\pp\mm};
\draw[fill] (9.5,0.75) circle [radius=0.05];
\node[above right] at (10.2,1.2) {$(S_2 \times S_6)$\,\mm\mm\mm};
\draw[fill] (10.2,1.2) circle [radius=0.05];
\node[above left] at (8.2,1.2) {$(S_2 \times S_5)$\,\mm\mm\pp};
\draw[fill] (8.2,1.2) circle [radius=0.05];
%%
\node[left] at (7.5,-0.75) {$(S_1 \times S_3)$\,\pp\pp\pp};
\draw[fill] (7.5,-0.75) circle [radius=0.05];
\node[right] at (9.5,-0.75) {$(S_1 \times S_4)$\,\pp\pp\mm};
\draw[fill] (9.5,-0.75) circle [radius=0.05];
\node[above right] at (10.2,-0.3) {$(S_1 \times S_6)$\,\pp\mm\mm};
\draw[fill] (10.2,-0.3) circle [radius=0.05];
\node[left] at (8.2,-0.3) {$(S_1 \times S_5)$\,\pp\mm\pp};
\draw[fill] (8.2,-0.3) circle [radius=0.05];
%%
\draw[-] (7.5,0.75) -- (9.5,0.75);
\draw[-] (9.5,0.75) -- (10.2,1.2);
\draw[-] (10.2,1.2) -- (8.2,1.2);
\draw[-] (8.2,1.2) -- (7.5,0.75);
%%
\draw[-] (7.5,-0.75) -- (9.5,-0.75);
\draw[-] (9.5,-0.75) -- (10.2,-0.3);
\draw[-,dashed] (10.2,-0.3) -- (8.2,-0.3);
\draw[-,dashed] (8.2,-0.3) -- (7.5,-0.75);
%%
\draw[-] (7.5,0.75) -- (7.5,-0.75);
\draw[-] (9.5,0.75) -- (9.5,-0.75);
\draw[-] (10.2,1.2) -- (10.2,-0.3);
\draw[-,dashed] (8.2,1.2) -- (8.2,-0.3);
\end{tikzpicture}
\caption{\label{multv}Visualization of multiplication of two cubical sets.}
\end{figure*}

@@

Formally, the denotation of types is as follows:
\[\begin{array}{rcl}
\den{0} &=& \emptyset \\
\den{1} &=& \{ \star \} \\
\den{\tau_1 + \tau_2} &=& \den{\tau_1} \oplus \den{\tau_2} \\
\den{\tau_1 * \tau_2} &=& \den{\tau_1} \otimes \den{\tau_2} \\
\den{- \tau} &=& \ominus \den{\tau} \\
\\
\noalign{\mbox{where:}\hfill}
\\
S_1 \oplus S_2 &=& S_1 \uplus S_2 \\
S \oplus (\nodet{\cubt_1}{\cubt_2}) &=& 
  (\nodet{S}{\phantom{S}}) \oplus (\nodet{\cubt_1}{\cubt_2}) \\
(\nodet{\cubt_1}{\cubt_2}) \oplus S &=& 
  (\nodet{\cubt_1}{\cubt_2}) \oplus (\nodet{S}{\phantom{S}}) \\
(\nodet{\cubt_1}{\cubt_2}) \oplus (\nodet{\cubt_3}{\cubt_4}) &=& 
  \nodet{\cubt_1 \oplus \cubt_3}{\cubt_2 \oplus \cubt_4} \\
\\
S_1 \otimes S_2 &=& S_1 \times S_2 \\
S \otimes (\nodet{\cubt_1}{\cubt_2}) &=& 
  \nodet{S \otimes \cubt_1}{S \otimes \cubt_2} \\
(\nodet{\cubt_1}{\cubt_2}) \otimes \cubt &=& 
  \nodet{\cubt_1 \otimes \cubt}{\cubt_2 \otimes \cubt} \\
\\
\ominus~S &=& \nodet{\phantom{S}}{S} \\
\ominus~\nodet{\cubt_1}{\cubt_2} &=& \nodet{\ominus~\cubt_2}{\ominus~\cubt_1} 
\end{array}\]
The type 0 maps to the empty set. The type 1 maps to a singleton set. The sum
of $0$-dimensional types is the disjoint union as expected. For cubes of
higher dimensions, the subspaces are recursively added. If the two types to
be added are of different dimensions, it is necessary to embed the smaller
type in the larger dimension using the fact (to be justified using a
$\Pi$-combinator that witnesses this isomorphism in the next section) that
$t$ is equivalent to $t-0$ and to $(t-0)-0$ and so on, i.e.,
\[
S = \nodet{S}{\phantom{S}} = 
  \nodet{\nodet{S}{\phantom{S}}}
        {\phantom{{\nodet{S}{\phantom{S}}}}} = \ldots \\
\]
The product of 0-dimensional types is the cartesian product of sets. The
definition of negation is natural: it recursively swaps the positive and
negative subspaces. We note that the general definitions of sums and products
reduce to the conventional sums and products for $0$-dimensional types. Also
the sum of 1-dimensional types reduces to the sum used in the \textbf{Int}
construction and that addition and negation do not increase the
dimensions. The interesting new development is the definition of
products. The example in Fig.~\ref{mult} illustrates the general pattern: the
product of an $n$-dimensional type with an $m$-dimensional type gives an
$(n+m)$-dimensional type. For low dimensions such as the ones used in the
example above, it is possible to visualize what is happening in coordinate
space (see Fig.~\ref{multv}).

@@

@@

Now that the type structure is defined, we turn our attention to the notion
of values. For conventional first-order types, the values are also
conventional:
\[
\infer{() : 1}{} 
\qquad
\infer{\inl{v} : t_1 + t_2}{v : t_1}
\qquad
\infer{\inr{v} : t_1 + t_2}{v : t_2}
\qquad
\infer{(v_1,v_2) : t_1 * t_2}{v_1 : t_1 & v_2 : t_2}
\]
For cubical types, the values are more involved. Intuitively, a value of the
$n$-dimensional type $\tau$ is an element of one of the sets located in a
corner of the $n$-dimensional cube denoted by $\tau$. Thus to specify a
value, we must specify one of the corners of the cube which can easily be
done using a sequence $p$ of $+$ and $-$ polarities indicating how to
navigate the cube in each successive dimension starting from a fixed origin
to reach the desired corner. We denote such values as $v^{p}$ and use
$\epsilon$ for the empty sequence of polarities. One can think of the
sequence of signs as denoting a path in the binary tree representation of
cubical types. Formally:
\[\begin{array}{c}
\infer[\textit{base}]{v^{+} : t}{v : t} \\
\\
\infer[\textit{left}]{(\inl{v})^{p} : \tau_1 + \tau_2}{v^{p} : \tau_1}
\qquad
\infer[\textit{right}]{(\inr{v})^{p} : \tau_1 + \tau_2}{v^{p} : \tau_2} \\
\\
\infer[\textit{prod}]{(v_1,v_2)^{p_1 \cdot p_2} : \tau_1 * \tau_2}
      {v_1^{p_1} : \tau_1 & v_2^{p_2} : \tau_2} \\
\\
\infer[\textit{neg}]{v^{\negp{p}} : - \tau}{v^p : \tau} 
\end{array}\]
The rule \textit{base} states that a conventional first-order value can be
viewed as trivial $0$-dimensional cubical value. The rules \textit{left} and
\textit{right} reflect the fact that sums do not increase the dimension and
works pointwise. The rule \textit{prod} is the most involved one: it
increases the dimension by \emph{concatenating} the two dimensions in its
arguments. For example, if we pair $v_1^{\pp}$ and $v_2^{\mm\pp}$ we get
$(v_1,v_2)^{\pp\mm\pp}$. The rule \textit{neg} shows that a negative value
$v$ is located at the ``opposite'' corner of $v$ which is calculated as
follows:
\[\begin{array}{rcl}
\negp{\epsilon} &=& \epsilon \\
\negp{+p} &=& -\negp{p} \\
\negp{-p} &=& +\negp{p}
\end{array}\]

%%%%%%%%%%%%%%%%%%
\subsection{Combinators}

>>>>

\newcommand{\bdim}[1]{\textsf{dim}(#1)}
\newcommand{\Rule}[4]{
\makebox{{\rm #1}
$\displaystyle
\frac{\begin{array}{l}#2\\\end{array}}
{\begin{array}{l}#3\\\end{array}}$
 #4}}
\newcommand{\proves}{\vdash}
\newcommand{\symc}[1]{\mathit{sym}~#1}
\newcommand{\jdg}[3]{#2 \proves_{#1} #3}
\newcommand{\adjoint}[1]{#1^{\dagger}}
\newcommand{\iso}{\leftrightarrow}
\newcommand{\identlp}{\mathit{identl}_+}
\newcommand{\identrp}{\mathit{identr}_+}
\newcommand{\swapp}{\mathit{swap}_+}
\newcommand{\assoclp}{\mathit{assocl}_+}
\newcommand{\assocrp}{\mathit{assocr}_+}
\newcommand{\identlt}{\mathit{identl}_*}
\newcommand{\identrt}{\mathit{identr}_*}
\newcommand{\swapt}{\mathit{swap}_*}
\newcommand{\assoclt}{\mathit{assocl}_*}
\newcommand{\assocrt}{\mathit{assocr}_*}
\newcommand{\distz}{\mathit{dist}_0}
\newcommand{\factorz}{\mathit{factor}_0}
\newcommand{\dist}{\mathit{dist}}
\newcommand{\factor}{\mathit{factor}}
\newcommand{\idc}{\mathit{id}}
\newcommand{\plus}{\raisebox{.4\height}{\scalebox{.6}{+}}}
\newcommand{\minus}{\raisebox{.4\height}{\scalebox{.8}{-}}}
\newcommand{\mm}{\texttt{-}}
\newcommand{\pp}{\texttt{+}}
\newcommand{\negp}[1]{\textit{neg}(#1)}
\newcommand{\inl}[1]{\textsf{inl}~#1}
\newcommand{\inr}[1]{\textsf{inr}~#1}
\newcommand{\lolli}{\multimap} 
\newcommand{\cubt}{\mathbb{T}}
\newcommand{\den}[1]{\llbracket #1 \rrbracket}
\newcommand{\nodet}[2]{\fcolorbox{black}{white}{$#1$}\fcolorbox{black}{gray!20}{$#2$}}
\newcommand{\hast}{$:\mkern -2.5mu:\;$}

@@

\begin{figure*}
\[\begin{array}{c}
\nodet{S_1}{S_2}
\quad\otimes\quad
\nodet{(\nodet{S_3}{S_4})}{(\nodet{S_5}{S_6})} \quad= \\
\\
\nodet{(\nodet{{(\nodet{S_1 \times S_3}{S_1 \times S_4})}}
              {{(\nodet{S_1 \times S_5}{S_1 \times S_6})}})}
      {(\nodet{{(\nodet{S_2 \times S_3}{S_2 \times S_4})}}
              {{(\nodet{S_2 \times S_5}{S_2 \times S_6})}})}
\end{array}\]

\begin{tikzpicture}
\node[above] at (0,0) {\pp};
\node[left] at (0,0) {$S_1$};
\draw[fill] (0,0) circle [radius=0.05];
\node[above] at (0.6,0) {\mm};
\node[right] at (0.6,0) {$S_2$};
\draw[fill] (0.6,0) circle [radius=0.05];
\draw[-] (0,0) -- (0.6,0);
\node at (1.6,0) {$\otimes$};

%%
\node[left] at (2.5,-0.5) {\pp\pp};
\node[below] at (2.5,-0.5) {$S_3$};
\draw[fill] (2.5,-0.5) circle [radius=0.05];
\node[right] at (3.5,-0.5) {\pp\mm};
\node[below] at (3.5,-0.5) {$S_4$};
\draw[fill] (3.5,-0.5) circle [radius=0.05];
\draw[-] (2.5,-0.5) -- (3.5,-0.5);
\draw[-] (2.5,-0.5) -- (2.5,0.5);
\node[left] at (2.5,0.5) {\mm\pp};
\node[above] at (2.5,0.5) {$S_5$};
\draw[fill] (2.5,0.5) circle [radius=0.05];
\node[right] at (3.5,0.5) {\mm\mm};
\node[above] at (3.5,0.5) {$S_6$};
\draw[fill] (3.5,0.5) circle [radius=0.05];
\draw[-] (2.5,0.5) -- (3.5,0.5);
\draw[-] (3.5,-0.5) -- (3.5,0.5);
%% 
\node at (5,0) {$=$};
%% 
\node[left] at (7.5,0.75) {$(S_2 \times S_3)$\,\mm\pp\pp};
\draw[fill] (7.5,0.75) circle [radius=0.05];
\node[right] at (9.5,0.75) {$(S_2 \times S_4)$\,\mm\pp\mm};
\draw[fill] (9.5,0.75) circle [radius=0.05];
\node[above right] at (10.2,1.2) {$(S_2 \times S_6)$\,\mm\mm\mm};
\draw[fill] (10.2,1.2) circle [radius=0.05];
\node[above left] at (8.2,1.2) {$(S_2 \times S_5)$\,\mm\mm\pp};
\draw[fill] (8.2,1.2) circle [radius=0.05];
%%
\node[left] at (7.5,-0.75) {$(S_1 \times S_3)$\,\pp\pp\pp};
\draw[fill] (7.5,-0.75) circle [radius=0.05];
\node[right] at (9.5,-0.75) {$(S_1 \times S_4)$\,\pp\pp\mm};
\draw[fill] (9.5,-0.75) circle [radius=0.05];
\node[above right] at (10.2,-0.3) {$(S_1 \times S_6)$\,\pp\mm\mm};
\draw[fill] (10.2,-0.3) circle [radius=0.05];
\node[left] at (8.2,-0.3) {$(S_1 \times S_5)$\,\pp\mm\pp};
\draw[fill] (8.2,-0.3) circle [radius=0.05];
%%
\draw[-] (7.5,0.75) -- (9.5,0.75);
\draw[-] (9.5,0.75) -- (10.2,1.2);
\draw[-] (10.2,1.2) -- (8.2,1.2);
\draw[-] (8.2,1.2) -- (7.5,0.75);
%%
\draw[-] (7.5,-0.75) -- (9.5,-0.75);
\draw[-] (9.5,-0.75) -- (10.2,-0.3);
\draw[-,dashed] (10.2,-0.3) -- (8.2,-0.3);
\draw[-,dashed] (8.2,-0.3) -- (7.5,-0.75);
%%
\draw[-] (7.5,0.75) -- (7.5,-0.75);
\draw[-] (9.5,0.75) -- (9.5,-0.75);
\draw[-] (10.2,1.2) -- (10.2,-0.3);
\draw[-,dashed] (8.2,1.2) -- (8.2,-0.3);
\end{tikzpicture}
\caption{\label{mult}Example of multiplication of two cubical sets.}
\end{figure*}

@@

Formally, the denotation of types is as follows:
\[\begin{array}{rcl}
\den{0} &=& \emptyset \\
\den{1} &=& \{ \star \} \\
\den{\tau_1 + \tau_2} &=& \den{\tau_1} \oplus \den{\tau_2} \\
\den{\tau_1 * \tau_2} &=& \den{\tau_1} \otimes \den{\tau_2} \\
\den{- \tau} &=& \ominus \den{\tau} \\
\\
\noalign{\mbox{where:}\hfill}
\\
S_1 \oplus S_2 &=& S_1 \uplus S_2 \\
S \oplus (\nodet{\cubt_1}{\cubt_2}) &=& \nodet{S \oplus \cubt_1}{\cubt_2} \\
(\nodet{\cubt_1}{\cubt_2}) \oplus S &=& \nodet{\cubt_1 \oplus S}{\cubt_2} \\
(\nodet{\cubt_1}{\cubt_2}) \oplus (\nodet{\cubt_3}{\cubt_4}) &=& 
  \nodet{\cubt_1 \oplus \cubt_3}{\cubt_2 \oplus \cubt_4} \\
\\
S_1 \otimes S_2 &=& S_1 \times S_2 \\
S \otimes (\nodet{\cubt_1}{\cubt_2}) &=& 
  \nodet{S \otimes \cubt_1}{S \otimes \cubt_2} \\
(\nodet{\cubt_1}{\cubt_2}) \otimes \cubt &=& 
  \nodet{\cubt_1 \otimes \cubt}{\cubt_2 \otimes \cubt} \\
\\
\ominus~S &=& \nodet{\phantom{S}}{S} \\
\ominus~\nodet{\cubt_1}{\cubt_2} &=& \nodet{\ominus~\cubt_2}{\ominus~\cubt_1} 
\end{array}\]
The type 0 maps to the empty set. The type 1 maps to a singleton set. The sum
of $0$-dimensional types is the disjoint union as expected. For cubes of
higher dimensions, the subspaces are recursively added. Note that the sum of
1-dimensional types reduces to the sum used in the \textbf{Int} construction.
The definition of negation is natural: it recursively swaps the positive and
negative subspaces. The product of 0-dimensional types is the cartesian
product of sets. For cubes of higher-dimensions $n$ and $m$, the result is of
dimension $(n+m)$. The example in Fig.~\ref{mult} illustrates the idea using
the product of 1-dimensional cube (i.e., a line) with a 2-dimensional cube
(i.e., a square). The result is a 3-dimensional cube as illustrated.

@@

\begin{verbatim}
A critical isomorphism that is 
not reflected at all in the semantics
is that 1-1 = 0

The denotation of 1-1 is a line;
the denotation of 0 is the empty set

[ {*} | {*} ] --> []

Something like the simplification
of Conway games here perhaps?

There are two values of type
1-1

() 
and 
()^-

epsilon should take one and map
to the other because it can
never produce something of type 0

In general, we have several direction
not just left and right like in the
Int construction. In 3D we have
8 directions: +++, ++-, ..., ---
We have an epsilon that 
takes ++- and flips it to --+
and another that flips
+++ to --- and so on

So we should have 8 interpreters
for 3D one for each direction!!!

The ++- interpreter would 
manipulates values like
(v1,v2,v3) and would apply
the normal combinators to v1, v2, 
and the adjoint to v3 and so on.
\end{verbatim}

@@

Now that the type structure is defined, we turn our attention to the notion
of values. Intuitively, a value of the $n$-dimensional type $\tau$ is an
element of one of the sets located in one of the corners of the
$n$-dimensional cube denoted by $\tau$. Thus to specify a value, we must
first specify one of the corners of the cube (or equivalently one of the
leaves in the binary tree representation) which can easily be done using a
sequence $p$ of $+$ and $-$ polarities indicating how to navigate the cube in
each successive dimension starting from a fixed origin to reach the desired
corner. We write $v^{p}$ for the value $v$ located at corner $p$ of the cube
associated with its type. We use $\epsilon$ for the empty sequence of
polarities and identify $v$ with $v^\epsilon$. Note that the polarities
doesn't completely specify the type since different types like $(1+(1+1))$
and $((1+1)+1)$ are assigned the same denotation. What the path $p$ specifies
is the \emph{polarity} of the value, or its ``orientation'' in the space
denoted by its type. Formally:
\[\begin{array}{c}
\infer{() : 1}{} \\
\qquad
\infer[\textit{neg}]{v^{\negp{p}} : - \tau}{v^p : \tau} 
\\
\infer[\textit{left}]{(\inl{v})^{p} : \tau_1 + \tau_2}{v^{p} : \tau_1}
\qquad
\infer[\textit{right}]{(\inr{v})^{p} : \tau_1 + \tau_2}{v^{p} : \tau_2} 
\\
\infer[\textit{prod}]{(v_1,v_2)^{p_1 \cdot p_2} : \tau_1 * \tau_2}
      {v_1^{p_1} : \tau_1 & v_2^{p_2} : \tau_2} 
\end{array}\]
The rules \textit{left} and \textit{right} reflect the fact that sums do not
increase the dimension. Note that when $p$ is $\epsilon$, we get the
conventional values for the 0-dimensional sum type. The rule \textit{prod} is
the most involved one: it increases the dimension by \emph{concatenating} the
two dimensions of its arguments. For example, if we pair $v_1^{\pp}$ and
$v_2^{\mm\pp}$ we get $(v_1,v_2)^{\pp\mm\pp}$. (See Fig.~\ref{mult} for the
illustration.) Note again that if both components are 0-dimensional, the pair
remains 0-dimensional and we recover the usual rule for typing values of
product types. The rule \textit{neg} uses the function below which states
that the negation of a value $v$ is the same value $v$ located at the
``opposite'' corner of the cube:
\[\begin{array}{rcl}
\negp{\epsilon} &=& \epsilon \\
\negp{+p} &=& -\negp{p} \\
\negp{-p} &=& +\negp{p}
\end{array}\]

\begin{verbatim}
We only have -
which flips ALL the directions simultaneously.
So from ++-+--+ you can only go to
        --+-++-
What if you wanted to go to
        ++++--+
I think we use the * functor to flip the 
direction we want. 
\end{verbatim}

%%%%%%%%%%%%%%%%%%
\subsection{Combinators: Example}

Consider the following simple function on 0-dimensional sum types:
\[\begin{array}{rcl}
\textsf{swapP} &\hast& (\tau_1 + \tau_2) \rightarrow (\tau_2 + \tau_1) \\
\textsf{swapP} (\inl{v}) &=& inr{v} \\
\textsf{swapP} (\inr{v}) &=& inl{v}
\end{code}
Given our setup, this just works $n$-dimensional types. And so do most of the
$\Pi$ combinators for that matter. Only $\eta$ and $\epsilon$ seem to need
thinking. How and why would $1-1$ which is a 1-dimensional line be the same
as the empty type which is a 0-dimensional thing. And how do we generalize
for arbitrary group identities at higher dimensions. We need a mechanism for
cubes with subspaces that ``cancel'' to map to equivalent smaller subcubes.

%%%%%%%%%%%%%%%%%%
\subsection{$\Pi$ Combinators}
